---
title: "tidyverse"
format: pdf
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Task 1
### Question A
```{r}
?read_csv()
```

**In 1-2 sentences, explain why we can not use specifically the read_csv() to read in these data.**

The two text files we are prompted to read into R do not contain comma separated values--they use semicolons instead. This is why we cannot use read_csv() and must instead use read_csv2(), due to the difference in delimiter.

```{r}
data <- read_csv2("Data/data.txt")

data
```

### Question B
```{r}
six <- read_delim("Data/data2.txt",
                  delim = "6",
                  col_types = "fdc")

six
```

## Task 2
### Question A
```{r}
trailblazer <- read_csv("Data/trailblazer.csv")

glimpse(trailblazer)
```

### Question B
```{r}
trailblazer_longer <- trailblazer |>
                        pivot_longer(cols = 2:11,
                          names_to = c("Game", "Location"),
                          names_sep = "_",
                          values_to = "Points")

head(trailblazer_longer, n=5)
```

### Question C
```{r}
trailblazer_wider <- trailblazer_longer |>
                       pivot_wider(names_from = Location,
                                   values_from = Points) |>
                       group_by(Player) |>
                       mutate(mean_home = mean(Home, na.rm = T),
                              mean_away = mean(Away, na.rm = T),
                              points_diff = mean_home - mean_away) |>
                       arrange(desc(points_diff))
  
print(trailblazer_wider, n = 90)
```

**In 1 sentence, state which players scored, on average, more points at home through the first 10 games of the season than away.**

On average, player Jusuf Nurkic scored more points at home through the first 10 games of the season than at away games.

## Task 3
### Question A
```{r}

```

### Question B
```{r}

```

## Task 4
```{r}

```